---
title: COVID visuals - daily new cases per million
author: admin
date: '2021-12-26'
slug: covid-visuals
categories:
  - R
tags:
  - plot
subtitle: ''
summary: ''
authors: []
lastmod: '2021-12-26T15:17:54+03:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache=FALSE, warning=FALSE,
                      message=FALSE, echo=TRUE, dpi=180,
                      fig.width=8, fig.height=6)
ggplot2::theme_set(ggplot2::theme_light())
```

# Load libraries

```{r}
library(tidyverse)
library(rvest)
library(slider)
```

## Input data - Worldometer



The data is scraped from a table in [Worldometer coronavirus site](https://www.worldometers.info/coronavirus/)

Why *yesterday*? Because, *today*'s data is incomplete, since not all countries update data at the same time.

```{r}
worldometer_url <- "https://www.worldometers.info/coronavirus/#countries"
covid_yesterday <- read_html(worldometer_url) %>% 
  html_node(xpath='//*[@id="main_table_countries_yesterday"]') %>%
  html_table() %>%
  as_tibble()
```

## Clean and analyze data

```{r}
covid_yesterday %>%
  janitor::clean_names() %>%
  filter(!is.na(number)) %>%
  select(country_other,new_cases,population) %>%
  mutate(across(c(new_cases,population),parse_number)) %>%
  top_n(20,new_cases) %>%
  mutate(new_per_1m=new_cases/population*100000, pop_m=round(population / 1000000, 2)) %>%
  select(-population) %>%
  arrange(-new_per_1m) %>% 
  head(10) %>% 
  kable(caption = "Top 10 countries with highest new cases per million")
```

## Another input data - John Hopkins Univ

```{r}
initial <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv") %>%
  janitor::clean_names()
```

Overall plot is quite crowded so I handpicked several countries.

```{r fig.width=6, fig.height=4}
select_countries <- c("Turkey","Brazil","US","Poland","Germany")

p <- initial %>%
  filter(country_region %in% select_countries) %>%
  select(country_region, starts_with("x")) %>%
  pivot_longer(-country_region, names_to="date", values_to="cases") %>%
  mutate(date=str_remove(date,"x")) %>%
  mutate(date=str_replace_all(date,"_","/")) %>%
  mutate(date=parse_date(date,"%m/%d/%y")) %>%
  mutate(new_case=cases-lag(cases)) %>%
  filter(!is.na(new_case), new_case > 0, new_case < 400000) %>%
  ggplot(aes(date,new_case, color=country_region)) +
  geom_line()

p
```

The plot is too rugged. Let's do weekly average by using excellent `slider::slide` function (instead of `zoo::rollmean`)

```{r}
initial %>%
  filter(country_region %in% select_countries) %>%
  select(country_region, starts_with("x")) %>%
  pivot_longer(-country_region, names_to="date", values_to="cases") %>%
  mutate(date=str_remove(date,"x")) %>%
  mutate(date=str_replace_all(date,"_","/")) %>%
  mutate(date=parse_date(date,"%m/%d/%y")) %>%
  mutate(new_case=cases-lag(cases)) %>%
  filter(!is.na(new_case), 
         new_case > 0, 
         new_case < 400000) %>% 
  mutate(new_case_roll= slide_dbl(new_case, 
                                  mean, 
                                  .before = 6)) %>% 
  ggplot(aes(date,new_case_roll, color=country_region)) +
  geom_line() -> p2

p2
```


using `plotly` package, we can have interactive results by issuing `ggplotly(p)`.

```{r fig.width=4, fig.height=3}
library(plotly)

ggplotly(p2) 
```

